Inferring neural networks from fluorescent calcium imaging using partial correlation and sample weighting

### 2014-05-22 article.tex
correlation does not imply causation

A common association measure is the Pearson correlation coefficient. 
\begin{align}
\rho_{ij} = \dfrac{cov(I,J)}{\sigma_I \sigma_J}
\end{align}
Even if \textit{correlation does not imply causation} [todo cite], it may give an insight on the underlying connectivity network. Nevertheless, the main challenge in inferring a network is to distinguish direct from indirect interactions [de la Fuente et al., 2004]. While the classical correlation coefficient characterize both direct and indirect interactions, the partial correlation coefficient quantifies association between variables conditionally on one or several other variables (the number of variables considered being the order of the partial correlation coefficient). [Shipley, 2002] suggests to consider correlation coefficients of all possible orders, i.e. conditioned on every subsets of the set of $p-2$ other variables, in order to take into account multiple indirect paths between a given pair of variables.  An obvious limitation to calculate high-order coefficient is the number of samples in conditioned data sets. Considering the large number of samples at our disposal in the connectomics challenge, we only consider the highest order of partial correlation coefficient. We thus try to estimate partial correlation of order $p-2$ $\rho_{ij | \mathcal{V}^{-i,j}}$. A way to find these coefficient [todo cite?] is to take the inverse of the covariance matrix to get the partial correlation matrix of order $p-2$, also known as the precision matrix [todo cite] .

>>>>

A common association measure is the Pearson correlation coefficient. 
\begin{align}
\rho_{ij} = \dfrac{cov(I,J)}{\sigma_I \sigma_J}
\end{align}
Even if \textit{correlation does not imply causation} [todo cite], it may give an insight on the underlying connectivity network. Nevertheless, the main challenge in inferring a network is to distinguish direct from indirect interactions [de la Fuente et al., 2004]. While the classical correlation coefficient characterize both direct and indirect interactions, the partial correlation coefficient quantifies association between variables conditionally on one or several other variables (the number of variables considered being the order of the partial correlation coefficient). [Shipley, 2002] suggests to consider correlation coefficients of all possible orders, i.e. conditioned on every subsets of the set of $p-2$ other variables, in order to take into account multiple indirect paths between a given pair of variables.  An obvious limitation to calculate high-order coefficient is the number of samples in conditioned data sets. Considering the large number of samples at our disposal in the connectomics challenge, we only consider the highest order of partial correlation coefficient. We thus try to estimate partial correlation matrix as the inferred network where a matrix element, denoted by $\rho_{ij | \mathcal{V}^{-i,j}}$, is the partial correlation coefficient of order $p-2$ between neurons $i$ and $j$. Note that this matrix is symmetric and thus it is impossible to determine the causal mechanism behind the interaction (I directly causes J, J directly causes I or both [de la Fuente, 2004])\footnote{Two other causal mechanisms may be implied by a non-zero partial coefficient. First, the pair of variables may be induced by a common (hidden) variable and second, I (respectively J) is conditionally correlated to a hidden variable affecting J (resp. I) [de la Fuente, 2004].} . A way to find these coefficient [todo cite?] is to take the inverse of the covariance matrix to get the partial correlation matrix of order $p-2$, also known as the precision matrix [todo cite] .
